---
title: "datafiles_for_ArcDyn_methods_paper"
author: "Douglas Yu"
date: "17/01/2019"
output: html_document
---
This code takes the output files from 5_merge_idx_meta_genomecov.Rmd and combines with the taxonomy data to create input files for the environmental and mock-sample analyses

setwd("~/Dropbox/Working_docs/Roslin_Greenland/2017/bulk_samples/ArcDyn_scripts/5_idxstats_tabulate")

```{r load packages}
library(tidyverse)  
library(readxl) # read excel spreadsheets
library(hablar) # data type conversion
library(arsenal) # utilities
```

This file provides the total number of reads per sample, used to calculate FST instead of FSL
```{r reformat seqkit stats output}
# A2B2 resequenced (most of) the same samples as AB, but fortunately, Earlham Institute coded the AB samples as IPO3916/7_A1 and the A2B2 samples as PlateA/B_A1. So I can use this column to join with the other metadata  
pathtofastqseqdata = "../.."  
 
# PlatesAB 
# sample identifier in soupdata has form IPO3916_A1, where IPO3916 is PlateA, IPO3917 is PlateB
fastq_read_counts_PlatesAB <- read_table(file.path( "../../../../2016/bulk_samples/PlatesAB_EI_20160512/fastq_read_counts_PlatesAB.txt"))

fastq_read_counts_PlatesAB <- fastq_read_counts_PlatesAB %>%
    tidyr::separate(file, into = c(NA, "gpfs", "home", "b042", "greenland", "plates", "platessub", "BWA", "sample", "fastqfile"), sep = "/", remove = TRUE) %>%
    mutate_at("sample", str_remove, "Sample_") %>% # remove "Sample_" from beginning of string
    dplyr::select(sample, tot_num_seqs = num_seqs) %>%
    dplyr::group_by(sample) %>%
    summarise(tot_num_seqs = sum(tot_num_seqs))

# PlatesA2B2
# sample identifier in soupdata has form PlateA_A1
fastq_read_counts_PlatesA2B2 <- read_table(file.path(pathtofastqseqdata, "PlatesA2B2/fastq_read_counts_PlatesA2B2.txt"))

fastq_read_counts_PlatesA2B2 <- fastq_read_counts_PlatesA2B2 %>%
    tidyr::separate(file, into = c(NA, "gpfs", "home", "b042", "greenland", "plates", "platessub", "BWA", "sample", "fastqfile"), sep = "/", remove = TRUE) %>%
    mutate_at("sample", str_remove, "Sample_PRO1322_") %>% # remove "Sample_PRO1322_" from beginning of string
    dplyr::select(sample, tot_num_seqs = num_seqs) %>%
    dplyr::group_by(sample) %>%
    summarise(tot_num_seqs = sum(tot_num_seqs))

# PlatesEF
fastq_read_counts_PlatesEF <- read_table(file.path(pathtofastqseqdata, "PlatesEF/fastq_read_counts_PlatesEF.txt"))

fastq_read_counts_PlatesEF <- fastq_read_counts_PlatesEF %>%
    tidyr::separate(file, into = c(NA, "gpfs", "home", "b042", "greenland", "plates", "platessub", "BWA", "sample", "fastqfile"), sep = "/", remove = TRUE) %>%
    mutate_at("sample", str_remove, "_[ACGT]{9}-[ACGT]{6}") %>% # remove index sequence from end of string, e.g. _TAGGTTAGG-GCCAAT
    dplyr::select(sample, tot_num_seqs = num_seqs) %>%
    dplyr::group_by(sample) %>%
    summarise(tot_num_seqs = sum(tot_num_seqs))

# PlatesGH
fastq_read_counts_PlatesGH <- read_table(file.path(pathtofastqseqdata, "PlatesGH/fastq_read_counts_PlatesGH.txt"))

fastq_read_counts_PlatesGH <- fastq_read_counts_PlatesGH %>%
    tidyr::separate(file, into = c(NA, "gpfs", "home", "b042", "greenland", "plates", "platessub", "BWA", "sample", "fastqfile"), sep = "/", remove = TRUE) %>%
    mutate_at("sample", str_remove, "Sample_PRO1747_") %>% # remove index sequence from end of string, Sample_PRO1747_
    dplyr::select(sample, tot_num_seqs = num_seqs) %>%
    dplyr::group_by(sample) %>%
    summarise(tot_num_seqs = sum(tot_num_seqs))

fastq_read_counts_PlatesABA2B2EFGH <- bind_rows(fastq_read_counts_PlatesAB, fastq_read_counts_PlatesA2B2, fastq_read_counts_PlatesEF, fastq_read_counts_PlatesGH)

# sanity check: should have: "IPO3916" "IPO3917" "PlateA"  "PlateB"  "PlateE"  "PlateF"  "PlateG"  "PlateH"
fastq_read_counts_PlatesABA2B2EFGH %>% 
    dplyr::select(sample) %>% 
    tidyr::separate(sample, sep = "_", into = c("sampleprefix", "samplesuffix")) %>%
    dplyr::count(sampleprefix) %>% 
    dplyr::arrange(sampleprefix)

rm(fastq_read_counts_PlatesAB, fastq_read_counts_PlatesA2B2, fastq_read_counts_PlatesEF, fastq_read_counts_PlatesGH) 
```

Load saved idx_meta_genomecov_ABA2B2EFGH files, barcodes and mitogenomes
```{r}
barcodesoutputidxstatstabulatefolder <- "../../PlatesABA2B2EFGH/minimap2_406barcodes_20190116" 
mitogenomesoutputidxstatstabulatefolder <- "../../PlatesABA2B2EFGH/minimap2_308mitogenomes_20190115"

idx_meta_genomecov_ABA2B2EFGH_barcodes <- read_delim(file.path(barcodesoutputidxstatstabulatefolder, "idx_meta_genomecov_ABA2B2EFGH_minimap2_20190116_406barcodes.txt"), delim = "\t", escape_double = FALSE, trim_ws = TRUE)

idx_meta_genomecov_ABA2B2EFGH_mitogenomes <- read_delim(file.path(mitogenomesoutputidxstatstabulatefolder, "idx_meta_genomecov_ABA2B2EFGH_minimap2_20190115_308mitogenomes.txt"), delim = "\t", escape_double = FALSE, trim_ws = TRUE)
```

Make table of barcode names after running the above for the barcode idx_meta_genomecov files
```{r}
# idx_meta_genomecov_ABA2B2EFGH_barcodes <- idx_meta_genomecov_ABA2B2EFGH
#  
# rm(idx_meta_genomecov_A2B2, idx_meta_genomecov_AB, idx_meta_genomecov_EF, idx_meta_genomecov_GH, idx_meta_genomecov_ABA2B2EFGH) 

# make species table from the Barcode names
barcodes <- idx_meta_genomecov_ABA2B2EFGH_barcodes %>% 
    dplyr::select(COI = mitogenome, COI_Species, COI_length = mt_length) %>% 
    dplyr::filter(COI_Species != "COI_Spike") %>% # remove COI_spike species 
    dplyr::select(-COI_Species) %>% # remove column indicating COI_spike species 
    distinct(COI, .keep_all = TRUE) %>% # keep only distinct values of mitogenome, should return 406 
    mutate(BOLD = COI) %>% # make new BOLD column with mitogenome 
    mutate(BOLD = str_extract(BOLD, "[A-Z]{3}[0-9]{4}$")) %>% # extract only the BOLD number
    mutate(BOLD = str_c("BOLD:", BOLD)) # add "BOLD:" as prefix. Have to do it this way because some of the BOLD numbers are not prefixed by "BOLD:"

# extract individual taxonomic ranks and re-merge Species and BOLD_ID. Species and BOLD_ID are messy because the epithets can include multiple elements
# cf_extrema_ACI8982, sp_AAF9271, tridactylus_ACH0172, cf_extrema_ACI8982, AAN6605_NA
# the AAN6605_NA is when there was no species epithet in the original name, not even "sp"
barcodes <- barcodes %>% 
    tidyr::separate(COI, c("Order", "Family", "Genus", "Species", "BOLD_ID"), sep = "_", remove = FALSE, extra = "merge") %>% 
    tidyr::unite("Species_BOLD", c("Species", "BOLD_ID"))

# the info line of the COI_Barcode fasta file is already well formatted: 
    # Diptera_Anthomyiidae_Zaphne_divisa_BOLD:AAG2441
    # Even this one can be parsed by first keeping together (extra = "merge") everything after the fourth column (tundrica_BOLD:AAG1723). then merge the Species and BOLD_ID  columns (frontata_tundrica_BOLD:AAG1723)
    # Diptera_Anthomyiidae_Zaphne_frontata_tundrica_BOLD:AAG1723
```

Make table of mitogenome names
```{r}
# idx_meta_genomecov_ABA2B2EFGH_mitogenomes <- idx_meta_genomecov_ABA2B2EFGH
#  
# rm(idx_meta_genomecov_A2B2, idx_meta_genomecov_AB, idx_meta_genomecov_EF, idx_meta_genomecov_GH, idx_meta_genomecov_ABA2B2EFGH) 

# taxonomyfilepath <- "~/Dropbox/Working_docs/Roslin_Greenland/2017/bulk_samples/ArcDyn_scripts/6_referenceseqs_metadata/mitogenomes_barcodes"
taxonomyfilepath <- "../6_referenceseqs_metadata/mitogenomes_barcodes"
taxonomyfile <- "Species_for_mitogenome_sequencing_20190123.xlsx"

mitotaxonomies <- read_excel(file.path(taxonomyfilepath, taxonomyfile), sheet = "mitogenome_database", na = "NA") %>% 
    dplyr::filter(COI_CytB_sanger_sequenced == "no" &  mitogenome_status != "failed") # 308
# mitogenome_status == "failed" indicates specimens that failed sequencing at Copenhagen and were then sent to EI. These species are given two rows. I only want the EI row
# COI_CytB_sanger_sequenced == "no" indicates specimens for which we did not do COI and cytB Sanger sequencing instead of mitogenome assembly (n = 41)

mitotaxonomies <- mitotaxonomies %>% 
    dplyr::select(Yinqiu_mitogenome_fastaname, Order, Family, Genus, Species_BOLD)

mitotaxonomies <- mitotaxonomies %>% 
    tidyr::separate(Species_BOLD, sep = " ", into = c("Genus2", "Species_BOLD")) %>% 
    dplyr::select(-Genus2)
# ifelse(mitotaxonomies$Genus == mitotaxonomies$Genus2, 1, 0) # Genus and Genus2 are the same, so i can delete Genus2

mitotaxonomies$Species_BOLD <- str_replace(mitotaxonomies$Species_BOLD, "_BOLD", "__BOLD") # delimit species epithet and BOLD:number

mitotaxonomies <- mitotaxonomies %>% 
    tidyr::separate(Species_BOLD, sep = "__", into = c("Species", "BOLD")) 
# separate species epithet and BOLD:number

# sanity check
mitotaxonomies %>% distinct(BOLD) %>% count() # 308 unique BOLD numbers

mitotaxonomies <-  left_join(idx_meta_genomecov_ABA2B2EFGH_mitogenomes, mitotaxonomies, by = c("mitogenome" = "Yinqiu_mitogenome_fastaname")) %>% 
    dplyr::filter(COI_Species != "COI_Spike") %>% # remove COI_spike species
    dplyr::select(-COI_Species) # remove column indicating COI_spike species

# names(mitotaxonomies)

mitotaxonomies <- mitotaxonomies %>% 
    dplyr::select(BOLD, Order, Family, Genus, mitogenome, mt_length) %>% 
    distinct(BOLD, .keep_all = TRUE) # merge and reduce to 308 unique mitogenomes
```

# species dataset #
Make species database by joining barcodes and mitotaxonomies
```{r}
names(barcodes) 
# [1] "COI"          "Order"        "Family"       "Genus"        "Species_BOLD" "COI_length"  
# [7] "BOLD"        
names(mitotaxonomies)
# [1] "BOLD"       "Order"      "Family"     "Genus"      "mitogenome" "mt_length" 

# The Order, Family, Genus, and Species names can be different between the barcode and mitogenome names (e.g. because of situations where the same BOLDID is given two provisional genus names, but only one is present in one of the databases). We use the BOLDID as the matching index, so we have some flexibility about which name to use. I am keeping the barcode Latin names (*.x) as this list is longer.
species <- left_join(barcodes, mitotaxonomies, by = c("BOLD" = "BOLD")) %>%
    dplyr::select(BOLD, Order = Order.x, Family = Family.x, Genus = Genus.x, Species_BOLD, mitogenome, COI, mt_length, COI_length)
    
# by = c("BOLD", "Order", "Family", "Genus"), 
species$sp <- seq(nrow(species))
species$sp <- str_c("sp", species$sp)

# mark some species for removal before environmental analyses
species <- species %>% 
    mutate(omit_from_env = case_when(
        Order %in% c("Sarcoptiformes", "Trombidiformes", "Entomobryomorpha", "Neelipleona", "Poduromorpha", "Symphypleona") ~ "Omit",
        TRUE ~ "Keep"
    )
)

species <- species %>% 
    dplyr::select(sp, everything(), omit_from_env)

# rm(barcodes, mitotaxonomies)

# example code to remove the omit_from_env species and the column
# species <- species %>% 
#     dplyr::filter(omit_from_env == "Keep") %>%  # remove the rows with these species
#     dplyr::select(-omit_from_env)   # remove this column
```

# spikes dataset #
```{r}
spikes <- idx_meta_genomecov_ABA2B2EFGH_barcodes %>% 
    dplyr::select(mitogenome, COI_Species) %>% 
    dplyr::filter(COI_Species == "COI_Spike") %>% 
    dplyr::select(mitogenome) %>% 
    distinct()
    
spikes$species <- seq(nrow(spikes)) 
spikes$species <- str_c("spike", spikes$species) 

spikes <- spikes %>% dplyr::select(species, full.name = mitogenome)
```

# env_design dataset #  Works for both barcodes and mitogenomes
```{r}
env_design <- idx_meta_genomecov_ABA2B2EFGH_barcodes %>% 
    dplyr::mutate(run = EI_RUN) %>%  
    dplyr::mutate(EI_RUN = str_remove(EI_RUN, "idx_meta_genomecov_")) %>%  
    dplyr::mutate(DateTrapRun = str_c(DateTrap, "_", EI_RUN)) %>% 
    dplyr::mutate(lysis_buffer_proportion = 1 / lysis_buffer_proportion) %>% # for downstream analysis
    dplyr::select(sample = DateTrapRun, date = Date, trap = Trap, run, lysis_buffer_proportion, mitogenome, COI_Species) %>% 
    dplyr::filter(COI_Species == "COI_Spike") %>% 
    dplyr::select(-COI_Species)

env_design <- env_design %>% 
    mutate(
    inputDNA = case_when(
        mitogenome == "Lepidoptera_Bombycidae_Bombyx_mori_COI" & run == "idx_meta_genomecov_AB" ~ 10,
        mitogenome == "Lepidoptera_Bombycidae_Bombyx_mori_COI" & run == "idx_meta_genomecov_A2B2" ~ 10,
        mitogenome == "Lepidoptera_Bombycidae_Bombyx_mori_COI" & run == "idx_meta_genomecov_EF" ~ 0.2,
        mitogenome == "Lepidoptera_Bombycidae_Bombyx_mori_COI" & run == "idx_meta_genomecov_GH" ~ 0.2
    )
)
env_design <- env_design %>% 
    mutate(
    inputDNA = case_when(
        mitogenome == "Coleoptera_Mordellidae_COI" & run == "idx_meta_genomecov_AB" ~ 20,
        mitogenome == "Coleoptera_Mordellidae_COI" & run == "idx_meta_genomecov_A2B2" ~ 20,
        mitogenome == "Coleoptera_Mordellidae_COI" & run == "idx_meta_genomecov_EF" ~ 0.4,
        mitogenome == "Coleoptera_Mordellidae_COI" & run == "idx_meta_genomecov_GH" ~ 0.4,
        TRUE ~ inputDNA
    )
)
env_design <- env_design %>% 
    mutate(
    inputDNA = case_when(
        mitogenome == "Coleoptera_Elateridae_COI" & run == "idx_meta_genomecov_AB" ~ 40,
        mitogenome == "Coleoptera_Elateridae_COI" & run == "idx_meta_genomecov_A2B2" ~ 40,
        mitogenome == "Coleoptera_Elateridae_COI" & run == "idx_meta_genomecov_EF" ~ 0.8,
        mitogenome == "Coleoptera_Elateridae_COI" & run == "idx_meta_genomecov_GH" ~ 0.8,
        TRUE ~ inputDNA
    )
)

env_design <- env_design %>% 
    tidyr::spread(mitogenome, inputDNA) %>% 
    dplyr::select(sample, date, trap, run, lysis_buffer_proportion, spike1 = Lepidoptera_Bombycidae_Bombyx_mori_COI, spike2 = Coleoptera_Mordellidae_COI, spike3 = Coleoptera_Elateridae_COI) %>% 
    arrange(run, date, trap)
 
names(env_design) 
```

# env_seq_depth dataset #  Works for both barcodes and mitogenomes
```{r}
env_seq_depth <- idx_meta_genomecov_ABA2B2EFGH_barcodes %>% 
    dplyr::mutate(run = EI_RUN) %>%  
    dplyr::mutate(EI_RUN = str_remove(EI_RUN, "idx_meta_genomecov_")) %>% 
    dplyr::mutate(DateTrapRun = str_c(DateTrap, "_", EI_RUN)) %>% 
    dplyr::select(DateTrapRun, sample) %>% 
    distinct(DateTrapRun, .keep_all = TRUE)

env_seq_depth <- left_join(env_seq_depth, fastq_read_counts_PlatesABA2B2EFGH) %>% 
    dplyr::select(-sample) %>% 
    dplyr::select(sample = DateTrapRun, mitogenome_seqs = tot_num_seqs) %>% 
    arrange(sample) 
```

# env_data_mitogenome dataset #
```{r}
env_data_mitogenome <- idx_meta_genomecov_ABA2B2EFGH_mitogenomes %>% 
    dplyr::mutate(run = EI_RUN) %>% 
    dplyr::mutate(EI_RUN = str_remove(EI_RUN, "idx_meta_genomecov_")) %>% 
    dplyr::mutate(DateTrapRun = str_c(DateTrap, "_", EI_RUN)) %>% 
    dplyr::select(DateTrapRun, mitogenome, PC, mapped_reads, COI_Species)

env_data_mitogenome <- left_join(env_data_mitogenome, species) # this will generate extra Order.y, Family.y, Genus.y columns from the species dataframe. We ignore them

env_data_mitogenome <- env_data_mitogenome %>% 
    dplyr::select(sample = DateTrapRun, sp, PC, mapped_reads, mitogenome, COI_Species)

spikes # table of COI_spike species, to refer to for the code below

# add spike1,2,3 to sp column
env_data_mitogenome <- env_data_mitogenome %>% 
    mutate(
    sp = case_when(
        mitogenome == "Lepidoptera_Bombycidae_Bombyx_mori_COI" ~ "spike1",
        mitogenome == "Coleoptera_Mordellidae_COI" ~ "spike2",
        mitogenome == "Coleoptera_Elateridae_COI" ~ "spike3",
        TRUE ~ as.character(sp)
    )
)

env_data_mitogenome <- env_data_mitogenome %>% 
    dplyr::filter(mapped_reads > 0) %>%  
    dplyr::select(sample, sp, PC, mapped_reads) 

env_data_mitogenome <- env_data_mitogenome %>% 
    mutate(
        PC = case_when(
            sp == "spike1" ~ NA_real_,
            sp == "spike2" ~ NA_real_,
            sp == "spike3" ~ NA_real_,
            TRUE ~ PC
        )
    )
```

# env_data_COI dataset #
```{r}
env_data_COI <- idx_meta_genomecov_ABA2B2EFGH_barcodes %>% 
    dplyr::mutate(run = EI_RUN) %>% 
    dplyr::mutate(EI_RUN = str_remove(EI_RUN, "idx_meta_genomecov_")) %>% 
    dplyr::mutate(DateTrapRun = str_c(DateTrap, "_", EI_RUN)) %>% 
    dplyr::select(DateTrapRun, mitogenome, PC, mapped_reads, COI_Species)

env_data_COI <- left_join(env_data_COI, species, by = c("mitogenome" = "COI")) # this will generate extra Order.y, Family.y, Genus.y columns from the species dataframe. We ignore them

env_data_COI <- env_data_COI %>% 
    dplyr::select(sample = DateTrapRun, sp, PC, mapped_reads, mitogenome, COI_Species)

spikes # table of COI_spike species, to refer to for the code below

# add spike1,2,3 to sp column
env_data_COI <- env_data_COI %>%  
    mutate(
    sp = case_when(
        mitogenome == "Lepidoptera_Bombycidae_Bombyx_mori_COI" ~ "spike1",
        mitogenome == "Coleoptera_Mordellidae_COI" ~ "spike2",
        mitogenome == "Coleoptera_Elateridae_COI" ~ "spike3",
        TRUE ~ as.character(sp)
    )
)

env_data_COI <- env_data_COI %>% 
    dplyr::filter(mapped_reads > 0) %>% 
    dplyr::select(sample, sp, PC, mapped_reads) 

env_data_COI <- env_data_COI %>% 
    mutate(
        PC = case_when(
            sp == "spike1" ~ NA_real_,
            sp == "spike2" ~ NA_real_,
            sp == "spike3" ~ NA_real_,
            TRUE ~ PC
        )
    )
```


Load saved mocks_idx_meta_genomecov files
```{r}
barcodesoutputidxstatstabulatefolder <- "../../PlatesABA2B2EFGH/minimap2_406barcodes_20190116" 
mitogenomesoutputidxstatstabulatefolder <- "../../PlatesABA2B2EFGH/minimap2_308mitogenomes_20190115"

mocks_idx_meta_genomecov_EFGH_barcodes <- read_delim(file.path(barcodesoutputidxstatstabulatefolder, "mocks_idx_meta_genomecov_EFGH_minimap2_20190116_406barcodes.txt"), delim = "\t", escape_double = FALSE, trim_ws = TRUE)

mocks_idx_meta_genomecov_EFGH_mitogenomes <- read_delim(file.path(mitogenomesoutputidxstatstabulatefolder, "mocks_idx_meta_genomecov_EFGH_minimap2_20190115_308mitogenomes.txt"), delim = "\t", escape_double = FALSE, trim_ws = TRUE)

mocks_idx_meta_genomecov_EFGH_barcodes <-  mocks_idx_meta_genomecov_EFGH_barcodes %>%  
    mutate(
        sample = case_when(
            sample == "PlateGH_COI_1_negative_control" ~ "PlateGH_negative_control_1",
            sample == "PlateGH_COI_2_negative_control" ~ "PlateGH_negative_control_2",
            TRUE ~ as.character(sample)
        )
    )

mocks_idx_meta_genomecov_EFGH_barcodes$sample %>% unique()

mocks_idx_meta_genomecov_EFGH_mitogenomes <-  mocks_idx_meta_genomecov_EFGH_mitogenomes %>% 
    mutate(
        sample = case_when(
            sample == "PlateGH_COI_1_negative_control" ~ "PlateGH_negative_control_1",
            sample == "PlateGH_COI_2_negative_control" ~ "PlateGH_negative_control_2",
            TRUE ~ as.character(sample)
        )
    )

mocks_idx_meta_genomecov_EFGH_mitogenomes$sample %>% unique() 
```

# mock_design # Works for both barcodes and mitogenomes
```{r}
mock_design <- left_join(mocks_idx_meta_genomecov_EFGH_barcodes, species, by = c("mitogenome" = "COI")) # I can use either the barcodes or mitogenomes dataset 

spikes # table of COI_spike species, to refer to for the code below

# add spike1,2,3 to sp column
mock_design <- mock_design %>% 
    mutate(
    sp = case_when(
        mitogenome == "Lepidoptera_Bombycidae_Bombyx_mori_COI" ~ "spike1",
        mitogenome == "Coleoptera_Mordellidae_COI" ~ "spike2",
        mitogenome == "Coleoptera_Elateridae_COI" ~ "spike3",
        TRUE ~ as.character(sp)
    )
)

mock_design <- mock_design %>% 
    dplyr::select(sample, plate, experiment, run, input_amount, input_sp = sp) %>% 
    dplyr::filter(input_amount > 0) %>% 
    arrange(experiment, plate, sample, run)
```

# mock_data_mitogenome #
```{r}
names(mocks_idx_meta_genomecov_EFGH_mitogenomes)
 
mock_data_mitogenome <- left_join(mocks_idx_meta_genomecov_EFGH_mitogenomes, species, by = c("mitogenome" = "mitogenome"))  

spikes # table of COI_spike species, to refer to for the code below

# add spike1,2,3 to sp column
mock_data_mitogenome <- mock_data_mitogenome %>% 
    mutate(
    sp = case_when(
        mitogenome == "Lepidoptera_Bombycidae_Bombyx_mori_COI" ~ "spike1",
        mitogenome == "Coleoptera_Mordellidae_COI" ~ "spike2",
        mitogenome == "Coleoptera_Elateridae_COI" ~ "spike3",
        TRUE ~ as.character(sp)
    )
)

mock_data_mitogenome <- mock_data_mitogenome %>% 
    dplyr::select(sample, plate, experiment, run, sp, PC, mapped_reads) %>% 
    arrange(experiment, plate, sample, run)
```

# mock_data_COI #
```{r}
names(mocks_idx_meta_genomecov_EFGH_barcodes)
 
mock_data_COI <- left_join(mocks_idx_meta_genomecov_EFGH_barcodes, species, by = c("mitogenome" = "COI")) 

spikes # table of COI_spike species, to refer to for the code below

# add spike1,2,3 to sp column
mock_data_COI <- mock_data_COI %>% 
    mutate(
    sp = case_when(
        mitogenome == "Lepidoptera_Bombycidae_Bombyx_mori_COI" ~ "spike1",
        mitogenome == "Coleoptera_Mordellidae_COI" ~ "spike2",
        mitogenome == "Coleoptera_Elateridae_COI" ~ "spike3",
        TRUE ~ as.character(sp)
    )
)

mock_data_COI <- mock_data_COI %>% 
    dplyr::select(sample, plate, experiment, run, sp, PC, mapped_reads) %>% 
    arrange(experiment, plate, sample, run)
```

# mock_seq_depth dataset #  Works for both barcodes and mitogenomes
```{r}
names(mocks_idx_meta_genomecov_EFGH_barcodes)

mock_seq_depth <- mocks_idx_meta_genomecov_EFGH_barcodes %>% 
    dplyr::select(sample, plate_well) %>% 
    distinct(sample, .keep_all = TRUE) %>% 
    arrange(sample)

mock_seq_depth <- left_join(mock_seq_depth, fastq_read_counts_PlatesABA2B2EFGH, by = c("plate_well" = "sample")) %>% 
    dplyr::select(-plate_well, mitogenome_seqs = tot_num_seqs) %>% 
    arrange(sample) 
```

Save datasets into a single RData file
```{r}
setwd("~/Dropbox/Working_docs/Roslin_Greenland/2018/manuscript_methods/20190124_Ecoletts/Rcode/Rdata")
getwd()

save(species, spikes, mock_design, mock_data_mitogenome, mock_data_COI,
     env_design, env_data_mitogenome, env_data_COI, mock_seq_depth, env_seq_depth,
     file = "input_data_step5_20190124.1.RData")
```

END

Loading input_data_step5_20190120.RData, which is correctly formatted for downstream statistical analysis. I use these datafiles as my template for the above code. For viewing convenience, I then put these files into an Excel spreadsheet (saved to the manuscript_methods/ folder)
```{r load input_data_step5_20190120.RData, eval=FALSE}
pathtoexampledata = "~/Dropbox/Working_docs/Roslin_Greenland/2018/manuscript_methods/EcoLetts_20190120/Rcode/Rdata"  

load(file.path(pathtoexampledata, "input_data_step5_20190120.RData")) # new version of data tables after re-running all steps of the pipeline, to confirm reproducibility and to make a cleaner dataset for uploading. 

tibble(env_data_COI)
tibble(env_data_mitogenome)
tibble(env_design)
tibble(env_seq_depth)  # tot_num_seqs for those samples
tibble(mock_data_COI)
tibble(mock_data_mitogenome)
tibble(mock_design)
tibble(mock_seq_depth)  # tot_num_seqs for those samples
tibble(species)
tibble(spikes)

# to make excel workbook so that i can see the target dataframes easily
# write_tsv(env_data_COI, "env_data_COI.tsv")
# write_tsv(env_data_mitogenome, "env_data_mitogenome.tsv")
# write_tsv(env_design, "env_design.tsv")
# write_tsv(env_seq_depth, "env_seq_depth.tsv")  # tot_num_seqs for those samples
# write_tsv(mock_data_COI, "mock_data_COI.tsv")
# write_tsv(mock_data_mitogenome, "mock_data_mitogenome.tsv")
# write_tsv(mock_design, "mock_design.tsv")
# write_tsv(mock_seq_depth, "mock_seq_depth.tsv")  # tot_num_seqs for those samples
# write_tsv(species, "species.tsv")
# write_tsv(spikes, "spikes.tsv")
```



Previous code. Ignore this

```{r soupdata308, eval=FALSE}
# mapping outputs 
soupdata <- read_tsv(file.path(pathto20190115data, "minimap2_308mitogenomes_20190115/idx_meta_genomecov_ABA2B2EFGH_minimap2_20190115_308mitogenomes.txt.zip"))

## sanity check:  should have: "IPO3916" "IPO3917" "PlateA"  "PlateB"  "PlateE"  "PlateF"  "PlateG"  "PlateH"
soupdata %>% 
    dplyr::select(sample) %>% 
    tidyr::separate(sample, sep = "_", into = c("sampleprefix", "samplesuffix")) %>% 
    dplyr::count(sampleprefix) %>% 
    arrange(sampleprefix)

names(soupdata)

#  [1] "EI_RUN"                                      "Date"                                       
#  [3] "Plot"                                        "Trap"                                       
#  [5] "Week"                                        "mitogenome"                                 
#  [7] "Order"                                       "Family"                                     
#  [9] "Genus"                                       "Species_BOLD"                               
# [11] "mapped_reads"                                "mapped_reads_COI_corr"                      
# [13] "mapped_reads_COI_lysis_corr"                 "sum_coverage"                               
# [15] "mean_coverage"                               "stddev"                                     
# [17] "coefvar"                                     "pct_coverage"                               
# [19] "samtools_filter"                             "mt_length"                                  
# [21] "lysis_buffer_orig_total_ul"                  "lysis_buffer_purified_ul"                   
# [23] "lysis_buffer_proportion"                     "Lepidoptera_Bombycidae_Bombyx_mori_10ng_COI"
# [25] "Coleoptera_Mordellidae_20ng_COI"             "Coleoptera_Elateridae_40ng_COI"             
# [27] "COI_corr_components"                         "COI_corr"                                   
# [29] "Concentration_ng_per_ul"                     "Sample_volume_ul"                           
# [31] "x.260_280_ratio"                             "sample"                                     
# [33] "ArcDyn_Plate_name"                           "EI_Plate_name"                              
# [35] "well"                                        "Sample_alias"                               
# [37] "TGAC_barcode"                                "pathname"                                   
# [39] "pathname_genomecov"                          "Full_name_of_the_sample"                    
# [41] "DateTrapEIRUN"                              



# mitogenomes_data
mitogenomes_data <- soupdata %>% dplyr::select(mitogenome, mt_length, Order, Family, Genus, Species_BOLD)

# env_design
## env_design, select sample, Date,
env_design <- soupdata %>% dplyr::select(sample = DateTrapEIRUN, date = Date, trap = Trap, run = EI_RUN, lysis_buffer_orig_total_ul, lysis_buffer_purified_ul, lysis_buffer_proportion)

## spike input DNA amount
## Plates AB & A2B2:  10, 20, 40 ng;  Plates EF & GH:  0.2, 0.4, 0.8 ng, for Bombycidae, Mordellidae, Elateridae, respectively
env_design <- env_design %>% mutate(
    Lepidoptera_Bombycidae_Bombyx_mori_COI = case_when(
        run == "idx_meta_genomecov_AB" ~ 10,
        run == "idx_meta_genomecov_A2B2" ~ 10,
        run == "idx_meta_genomecov_EF" ~ 0.2,
        run == "idx_meta_genomecov_GH" ~ 0.2
    )
)
env_design <- env_design %>% mutate(
    Coleoptera_Mordellidae_COI = case_when(
        run == "idx_meta_genomecov_AB" ~ 20,
        run == "idx_meta_genomecov_A2B2" ~ 20,
        run == "idx_meta_genomecov_EF" ~ 0.4,
        run == "idx_meta_genomecov_GH" ~ 0.4
    )
)
env_design <- env_design %>% mutate(
    Coleoptera_Elateridae_COI = case_when(
        run == "idx_meta_genomecov_AB" ~ 40,
        run == "idx_meta_genomecov_A2B2" ~ 40,
        run == "idx_meta_genomecov_EF" ~ 0.8,
        run == "idx_meta_genomecov_GH" ~ 0.8
    )
)

# summarise env_design
# env_design_summarised <- env_design %>%
#     dplyr::group_by(sample) %>%
#     dplyr::summarise(
#         date = first(date),
#         trap = first(trap),
#         run = first(run),
#         lysis_buffer_orig_total_ul = first(lysis_buffer_orig_total_ul),
#         lysis_buffer_purified_ul = first(lysis_buffer_purified_ul),
#         lysis_buffer_proportion = first(lysis_buffer_proportion),
#         Lepidoptera_Bombycidae_Bombyx_mori_COI = first(Lepidoptera_Bombycidae_Bombyx_mori_COI),
#         Coleoptera_Mordellidae_COI = first(Coleoptera_Mordellidae_COI),
#         Coleoptera_Elateridae_COI = first(Coleoptera_Elateridae_COI)
#     )

# env_data
## make long-format dataset of the COI_spike mapped reads, one row per sample and spike_species
## change names of the COI spike mapped reads to omit the ng portion
env_spikes <- soupdata %>%
    dplyr::select(sample = DateTrapEIRUN,
                  Lepidoptera_Bombycidae_Bombyx_mori_COI = Lepidoptera_Bombycidae_Bombyx_mori_10ng_COI,
                  Coleoptera_Mordellidae_COI = Coleoptera_Mordellidae_20ng_COI,
                  Coleoptera_Elateridae_COI = Coleoptera_Elateridae_40ng_COI,
                  sampleorig = sample) %>%
    tidyr::gather(species, mapped_reads_mitogenome,
                  Lepidoptera_Bombycidae_Bombyx_mori_COI,
                  Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    dplyr::distinct(sample, species, mapped_reads_mitogenome, sampleorig)

# extract mapped reads and pct_coverage for each mitogenome in each sample
env_data <- soupdata %>%
    dplyr::select(sample = DateTrapEIRUN, species = mitogenome, mapped_reads_mitogenome = mapped_reads, pct_coverage_mitogenome = pct_coverage, sampleorig = sample)

# bind rows with env_spikes dataset, with different cols for mapped reads to mitogenomes and COI spike
env_data <- bind_rows(env_data, env_spikes) %>% arrange(sample)

# add fastq read counts to dataset,
env_data <- left_join(env_data, fastq_read_counts_PlatesABA2B2EFGH, by = c("sampleorig" = "sample")) %>% dplyr::select(-sampleorig)

# create a column indicating mitogenome or COI_spike species, change column order
env_data$mito_spike <- if_else(grepl("_COI$", env_data$species), "spike", "mitogenome")
env_data <- env_data %>% dplyr::select(sample, mito_spike, everything())

# save(mitogenomes_data, env_design, env_data, file = "~/Dropbox/Working_docs/Roslin_Greenland/2018/manuscript_methods/20190109/Rcode/Rdata/arcdyn_suppfiles_soups_20190115_308mitogenomes.RData")
# loadfile syntax
# load("~/Dropbox/Working_docs/Roslin_Greenland/2018/manuscript_methods/20190109/Rcode/Rdata/arcdyn_suppfiles_soups_20190115_308mitogenomes.RData")
```

```{r mock data files for Otso on 20181221 using 308 mitogenomes, eval=FALSE}
# mock datasets 
setwd("~/Dropbox/Working_docs/Roslin_Greenland/2017/bulk_samples/PlatesABA2B2EFGH") 

mockEF <- read_excel("minimap2_308mitogenomes_20190115/mocks_idx_meta_genomecov_PlatesEF_20190115_308mitogenomes.xlsx", sheet = "mocks_idx_meta_genomecov_Plates", na = "NA")
mockEF <- hablar::convert(mockEF, num(Replicate, inputDNA_ng))
# mockEF$Replicate <- as.character(mockEF$Replicate)
mockGH <- read_excel("minimap2_308mitogenomes_20190115/mocks_idx_meta_genomecov_PlatesGH_20190115_308mitogenomes.xlsx", sheet = "mocks_idx_meta_genomecov_Plates", na = "NA")
mockGH <- hablar::convert(mockGH, num(Replicate, inputDNA_ng))

# old location
# mockEF <- read_excel("PlatesEF/EF_outputs_20190115_F2308_f0x2_q48_minimap2_308mitogenomes/mocks_idx_meta_genomecov_PlatesEF_20190115_308mitogenomes.xlsx", sheet = "mocks_idx_meta_genomecov_Plates", na = "NA")
# mockEF$Replicate <- as.character(mockEF$Replicate)
# mockGH <- read_excel("PlatesGH/GH_outputs_20190115_F2308_f0x2_q48_minimap2_308mitogenomes/mocks_idx_meta_genomecov_PlatesGH_20190115_308mitogenomes.xlsx", sheet = "mocks_idx_meta_genomecov_Plates", na = "NA")

mockall <- dplyr::bind_rows(mockEF, mockGH)

# add fastq read counts to dataset,
mockall <- left_join(mockall, fastq_read_counts_PlatesABA2B2EFGH, by = c("sample" = "sample"))

## to make mock_inputspp
### extract COI_spike columns, make a long-format dataset (with gather), change column name to mitogenome, reduce to unique COI_spike names, and add a new column with index numbers (1,2,3). Change COI_spike colnames to omit 10/20/40ng from name
mock_spikes <- mockall %>%
    dplyr::select(Lepidoptera_Bombycidae_Bombyx_mori_COI = Lepidoptera_Bombycidae_Bombyx_mori_10ng_COI, Coleoptera_Mordellidae_COI = Coleoptera_Mordellidae_20ng_COI, Coleoptera_Elateridae_COI = Coleoptera_Elateridae_40ng_COI) %>%
    tidyr::gather(species, mapped_reads, Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    dplyr::select(mitogenome = species) %>%
    dplyr::distinct(mitogenome)

mock_spikes <- cbind(mock_spikes, order = seq(nrow(mock_spikes)))

### extract mitogenome column, reduce to unique mitogenome names, and add a new column with index numbers (1,...,308)
mock_mitogenomes <- mockall %>%
    dplyr::select(mitogenome, mt_length) %>%
    dplyr::distinct(mitogenome, mt_length)

mock_mitogenomes <- cbind(mock_mitogenomes, order = seq(nrow(mock_mitogenomes)))

### bind mock_mitogenomes and mock_spikes, add spike or sp as prefix, merge with index number, change column order
mock_inputspp <- dplyr::bind_rows(mock_mitogenomes, mock_spikes)
mock_inputspp$species <- if_else(grepl("_COI$", mock_inputspp$mitogenome), "spike", "sp")
mock_inputspp <- mock_inputspp %>%
    unite(species, c("species", "order"), sep = "") %>%
    dplyr::select(species, mitogenome, mt_length)


## to make mock_design
## sample	experiment	run	lysis	input species	input mount
### select columns except for the COI_spike columns
mock_design <- mockall %>%
    dplyr::select(plate = Plate, shortname = Expt_shortname, experiment, run = Replicate, mitogenome_spike = mitogenome, input_amount = inputDNA_ng)
### left_join mock_inputspp. Do not include lysis (not necessary), do include mitogenome name
mock_design <- dplyr::left_join(mock_design, mock_inputspp, by = c("mitogenome_spike" = "mitogenome")) %>%
    dplyr::select(plate, shortname, experiment, run, mitogenome_spike, input_species = species, input_amount)

### select COI_spike species
mock_spikes <- mockall %>% dplyr::select(plate = Plate, shortname = Expt_shortname, experiment, run = Replicate, Lepidoptera_Bombycidae_Bombyx_mori_COI = Lepidoptera_Bombycidae_Bombyx_mori_10ng_COI, Coleoptera_Mordellidae_COI = Coleoptera_Mordellidae_20ng_COI, Coleoptera_Elateridae_COI = Coleoptera_Elateridae_40ng_COI) %>%
    tidyr::gather(mitogenome_spike, mapped_reads_COI, Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    dplyr::distinct(plate, shortname, experiment, run, mitogenome_spike, mapped_reads_COI)

### add input DNA amounts (since only PlatesEF & GH, the amts are 0.2, 0.4, 0.8 only)
mock_spikes <- mock_spikes %>% 
    mutate(input_amount = case_when(
        mitogenome_spike == "Lepidoptera_Bombycidae_Bombyx_mori_COI" ~ 0.2,
        mitogenome_spike == "Coleoptera_Mordellidae_COI" ~ 0.4,
        mitogenome_spike == "Coleoptera_Elateridae_COI" ~ 0.8
    )
)

mock_spikes <- left_join(mock_spikes, mock_inputspp, by = c("mitogenome_spike" = "mitogenome")) # %>% tidyr::separate(sample, c(NA, "Run"), sep = "_", remove = FALSE) #

mock_design$input_amount <- as.numeric(mock_design$input_amount)

mock_spikes <- mock_spikes %>% 
    dplyr::select(plate, shortname, experiment, run, mitogenome_spike, input_species = species, input_amount)

mock_design <- bind_rows(mock_design, mock_spikes) %>% 
    arrange(plate, shortname, run) # 4976 obs

mock_design <- mock_design %>% 
    unite("sample", c(plate, shortname, run), remove = FALSE)

names(mock_design)


## to make mock_data
## sample mitogenome_spike mapped_reads_mitogenome pct_coverage_mitogenome mapped_reads_COI

### select columns except for the COI_spike columns
mock_data <- mockall %>% 
    dplyr::select(plate = Plate, shortname = Expt_shortname, experiment, run = Replicate, sample, mitogenome_spike = mitogenome, mapped_reads_mitogenome = mapped_reads, pct_coverage_mitogenome = pct_coverage)

mock_spikes <- mockall %>% 
    dplyr::select(plate = Plate, shortname = Expt_shortname, experiment, run = Replicate, sample, Lepidoptera_Bombycidae_Bombyx_mori_COI = Lepidoptera_Bombycidae_Bombyx_mori_10ng_COI, Coleoptera_Mordellidae_COI = Coleoptera_Mordellidae_20ng_COI, Coleoptera_Elateridae_COI = Coleoptera_Elateridae_40ng_COI, tot_num_seqs) %>%
    tidyr::gather(mitogenome_spike, mapped_reads_mitogenome, Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    dplyr::distinct(plate, shortname, experiment, run, sample, mitogenome_spike, mapped_reads_mitogenome)

# names(mock_data)
# names(mock_spikes)

# bind mock_data and mock_spikes
mock_data <- bind_rows(mock_data, mock_spikes) %>% arrange(plate, shortname, run)

# add fastq read counts to dataset,
mock_data <- left_join(mock_data, fastq_read_counts_PlatesABA2B2EFGH, by = c("sample" = "sample")) %>% dplyr::select(-sample) # remove sample column with Plate_well info

mock_data <- mock_data %>% unite("sample", c(plate, shortname, run), remove = FALSE) # create new sample column using (plate, shortname, run). NB R automatically replaces old sample col with this new sample col
names(mock_data)


## save data files
# save(mock_inputspp, mock_design, mock_data, file = "~/Dropbox/Working_docs/Roslin_Greenland/2018/manuscript_methods/20190109/Rcode/Rdata/arcdyn_suppfiles_mocks_20190115_308mitogenomes.RData")

# loadfile syntax
# load("~/Dropbox/Working_docs/Roslin_Greenland/2018/manuscript_methods/20190109/Rcode/Rdata/arcdyn_suppfiles_mocks_20190115_308mitogenomes.RData")
```

```{r sample data files for Otso on 20181221 using 406 COI barcodes, eval=FALSE}
setwd("~/Dropbox/Working_docs/Roslin_Greenland/2017/bulk_samples/PlatesABA2B2EFGH") 
getwd()

# COI_barcode soups dataset
soupdata <- read_tsv("minimap2_406barcodes_20190116/idx_meta_genomecov_ABA2B2EFGH_minimap2_20190116_406barcodes.txt.zip")
# soupdata <- read_tsv("send_to_ArcDyn_collaborators_20180727_COI_Barcodes_only/idx_meta_genomecov_ABA2B2EFGH_minimap2_COIBarcodesOnly_20180727.txt.zip")

# mitogenomes_data
mitogenomes_data <- soupdata %>% dplyr::select(mitogenome, mt_length, Order, Family, Genus, Species_BOLD)

# env_design
## env_design, select sample, Date,
env_design <- soupdata %>% dplyr::select(sample = DateTrapEIRUN, date = Date, trap = Trap, run = EI_RUN, lysis_buffer_orig_total_ul, lysis_buffer_purified_ul, lysis_buffer_proportion)

## spike input DNA amount
## Plates AB & A2B2:  10, 20, 40 ng;  Plates EF & GH:  0.2, 0.4, 0.8 ng, for Bombycidae, Mordellidae, Elateridae, respectively
env_design <- env_design %>% mutate(
    Lepidoptera_Bombycidae_Bombyx_mori_COI = case_when(
        run == "idx_meta_genomecov_AB" ~ 10,
        run == "idx_meta_genomecov_A2B2" ~ 10,
        run == "idx_meta_genomecov_EF" ~ 0.2,
        run == "idx_meta_genomecov_GH" ~ 0.2
    )
)
env_design <- env_design %>% mutate(
    Coleoptera_Mordellidae_COI = case_when(
        run == "idx_meta_genomecov_AB" ~ 20,
        run == "idx_meta_genomecov_A2B2" ~ 20,
        run == "idx_meta_genomecov_EF" ~ 0.4,
        run == "idx_meta_genomecov_GH" ~ 0.4
    )
)
env_design <- env_design %>% mutate(
    Coleoptera_Elateridae_COI = case_when(
        run == "idx_meta_genomecov_AB" ~ 40,
        run == "idx_meta_genomecov_A2B2" ~ 40,
        run == "idx_meta_genomecov_EF" ~ 0.8,
        run == "idx_meta_genomecov_GH" ~ 0.8
    )
)

names(env_design)
# summarise env_design
# env_design_summarised <- env_design %>%
#     dplyr::group_by(sample) %>%
#     dplyr::summarise(
#         date = first(date),
#         trap = first(trap),
#         run = first(run),
#         lysis_buffer_orig_total_ul = first(lysis_buffer_orig_total_ul),
#         lysis_buffer_purified_ul = first(lysis_buffer_purified_ul),
#         lysis_buffer_proportion = first(lysis_buffer_proportion),
#         Lepidoptera_Bombycidae_Bombyx_mori_COI = first(Lepidoptera_Bombycidae_Bombyx_mori_COI),
#         Coleoptera_Mordellidae_COI = first(Coleoptera_Mordellidae_COI),
#         Coleoptera_Elateridae_COI = first(Coleoptera_Elateridae_COI)
#     )

names(soupdata)
# env_data
## make long-format dataset of the COI_spike mapped reads, one row per sample and spike_species
## change names of the COI spike mapped reads to omit the ng portion
env_spikes <- soupdata %>% dplyr::select(sample = DateTrapEIRUN, Lepidoptera_Bombycidae_Bombyx_mori_COI = Lepidoptera_Bombycidae_Bombyx_mori_10ng_COI, Coleoptera_Mordellidae_COI = Coleoptera_Mordellidae_20ng_COI, Coleoptera_Elateridae_COI = Coleoptera_Elateridae_40ng_COI, sampleorig = sample) %>%
    tidyr::gather(species, mapped_reads_mitogenome, Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    dplyr::distinct(sample, species, mapped_reads_mitogenome, sampleorig)

# extract mapped reads and pct_coverage for each mitogenome in each sample
env_data <- soupdata %>%
    dplyr::select(sample = DateTrapEIRUN, species = mitogenome, mapped_reads_mitogenome = mapped_reads, pct_coverage_mitogenome = pct_coverage, sampleorig = sample)

# bind rows with env_spikes dataset, with different cols for mapped reads to mitogenomes and COI spike
env_data <- bind_rows(env_data, env_spikes) %>% arrange(sample)

# add fastq read counts to dataset,
env_data <- left_join(env_data, fastq_read_counts_PlatesABA2B2EFGH, by = c("sampleorig" = "sample")) %>% dplyr::select(-sampleorig)

# create a column indicating mitogenome or COI_spike species, change column order
env_data$mito_spike <- if_else(grepl("_COI$", env_data$species), "spike", "mitogenome")
env_data <- env_data %>% dplyr::select(sample, mito_spike, everything())

names(env_data)

# save(mitogenomes_data, env_design, env_data, file = "~/Dropbox/Working_docs/Roslin_Greenland/2018/manuscript_methods/20190109/Rcode/Rdata/arcdyn_suppfiles_soups_20190116_406barcodes.RData")

```

```{r mock data files for Otso on 20181221 using 406 COI barcodes, eval=FALSE}
# mock datasets
## mockall
### read EF and GH mock datasets, select the same columns and bind_rows
setwd("~/Dropbox/Working_docs/Roslin_Greenland/2017/bulk_samples/PlatesABA2B2EFGH")


mockEF <- read_excel("minimap2_406barcodes_20190116/mocks_idx_meta_genomecov_PlatesEF_20190116_406barcodes.xlsx", sheet = "mocks_idx_meta_genomecov_Plates", na = "NA")
mockEF <- hablar::convert(mockEF, num(Replicate, inputDNA_ng))
mockGH <- read_excel("minimap2_406barcodes_20190116/mocks_idx_meta_genomecov_PlatesGH_20190116_406barcodes.xlsx", sheet = "mocks_idx_meta_genomecov_Plates", na = "NA")
mockGH <- hablar::convert(mockGH, num(Replicate, inputDNA_ng))

# mockEF <- read_excel("PlatesEF/EF_outputs_20190116_F2308_f0x2_q48_minimap2_406barcodes/mocks_idx_meta_genomecov_PlatesEF_20190116_406barcodes.xlsx", sheet = "mocks_idx_meta_genomecov_Plates", na = "NA") 
# mockEF <- hablar::convert(mockEF, num(Replicate, inputDNA_ng))
# mockGH <- read_excel("PlatesGH/GH_outputs_20190116_F2308_f0x2_q48_minimap2_406barcodes/mocks_idx_meta_genomecov_PlatesGH_20190116_406barcodes.xlsx", sheet = "mocks_idx_meta_genomecov_Plates", na = "NA")
# mockGH <- hablar::convert(mockGH, num(Replicate, inputDNA_ng))

mockall <- bind_rows(mockEF, mockGH)
names(mockall)

mockall <- mockall %>% dplyr::select(plate = Plate, shortname = Expt_shortname, experiment, run = Replicate, sample, mitogenome_spike = mitogenome, length = mt_length, mapped_reads_mitogenome = mapped_reads, pct_coverage_mitogenome = pct_coverage, input_amount = inputDNA_ng, Lepidoptera_Bombycidae_Bombyx_mori_COI = Lepidoptera_Bombycidae_Bombyx_mori_10ng_COI, Coleoptera_Mordellidae_COI = Coleoptera_Mordellidae_20ng_COI, Coleoptera_Elateridae_COI = Coleoptera_Elateridae_40ng_COI)
names(mockall)

## to make mock_inputspp
### extract COI_spike columns, make a long-format dataset (with gather), change column name to mitogenome, reduce to unique COI_spike names, and add a new column with index numbers (1,2,3). Change COI_spike colnames to omit 10/20/40ng from name
mock_spikes <- mockall %>%
    dplyr::select(Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    tidyr::gather(species, mapped_reads, Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    dplyr::select(mitogenome_spike = species) %>%
    dplyr::distinct(mitogenome_spike)
mock_spikes <- cbind(mock_spikes, order = seq(nrow(mock_spikes)))

### extract mitogenome column (which is the name of the COI barcode seqs), reduce to unique names, and add a new column with index numbers (1,...,406)
mock_mitogenomes <- mockall %>%
    dplyr::select(mitogenome_spike, length) %>%
    dplyr::distinct(mitogenome_spike, length)
mock_mitogenomes <- cbind(mock_mitogenomes, order = seq(nrow(mock_mitogenomes)))

### bind mock_mitogenomes and mock_spikes, add spike or sp as prefix, merge with index number, change column order
mock_inputspp <- dplyr::bind_rows(mock_mitogenomes, mock_spikes)
mock_inputspp$species <- if_else(grepl("_COI$", mock_inputspp$mitogenome_spike), "spike", "sp")
mock_inputspp <- mock_inputspp %>%
    unite(species, c("species", "order"), sep = "") %>%
    dplyr::select(species, mitogenome_spike, length)


## to make mock_design
## sample	experiment	run	lysis	input species	input mount
### select columns except for the COI_spike columns
mock_design <- mockall %>%
    dplyr::select(plate, shortname, experiment, run, mitogenome_spike, input_amount)
### left_join mock_inputspp. Do not include lysis (not necessary), do include mitogenome name
mock_design <- dplyr::left_join(mock_design, mock_inputspp, by = c("mitogenome_spike" = "mitogenome_spike")) %>%
    dplyr::select(plate, shortname, experiment, run, mitogenome_spike, input_species = species, input_amount)

### select COI_spike species
mock_spikes <- mockall %>% dplyr::select(plate, shortname, experiment, run, Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    tidyr::gather(mitogenome_spike, mapped_reads_COI, Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    dplyr::distinct(plate, shortname, experiment, run, mitogenome_spike, mapped_reads_COI)

### add input DNA amounts (since only PlatesEF & GH, the amts are 0.2, 0.4, 0.8 only)
mock_spikes <- mock_spikes %>% mutate(
    input_amount = case_when(
        mitogenome_spike == "Lepidoptera_Bombycidae_Bombyx_mori_COI" ~ 0.2,
        mitogenome_spike == "Coleoptera_Mordellidae_COI" ~ 0.4,
        mitogenome_spike == "Coleoptera_Elateridae_COI" ~ 0.8
    )
)

mock_spikes <- left_join(mock_spikes, mock_inputspp, by = c("mitogenome_spike" = "mitogenome_spike")) # %>% tidyr::separate(sample, c(NA, "Run"), sep = "_", remove = FALSE) #

mock_spikes <- mock_spikes %>% dplyr::select(plate, shortname, experiment, run, mitogenome_spike, input_species = species, input_amount)

mock_design <- bind_rows(mock_design, mock_spikes) %>% arrange(plate, shortname, run) # 6544 obs
mock_design <- mock_design %>% unite("sample", c(plate, shortname, run), remove = FALSE)
names(mock_design)



## to make mock_data
## sample mitogenome_spike mapped_reads_mitogenome pct_coverage_mitogenome mapped_reads_COI

### select columns except for the COI_spike columns
mock_data <- mockall %>% dplyr::select(plate, shortname, experiment, run, sample, mitogenome_spike, mapped_reads_mitogenome, pct_coverage_mitogenome)

mock_spikes <- mockall %>% dplyr::select(plate, shortname, experiment, run, sample,  Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    tidyr::gather(mitogenome_spike, mapped_reads_mitogenome, Lepidoptera_Bombycidae_Bombyx_mori_COI, Coleoptera_Mordellidae_COI, Coleoptera_Elateridae_COI) %>%
    dplyr::distinct(plate, shortname, experiment, run, sample, mitogenome_spike, mapped_reads_mitogenome)

names(mock_data)
names(mock_spikes)

# bind rows
mock_data <- bind_rows(mock_data, mock_spikes) %>% arrange(plate, shortname, run)

# add fastq read counts to dataset,
mock_data <- left_join(mock_data, fastq_read_counts_PlatesABA2B2EFGH, by = c("sample" = "sample")) %>% dplyr::select(-sample) # remove sample column with Plate_well info

mock_data <- mock_data %>% unite("sample", c(plate, shortname, run), remove = FALSE)

# create a column indicating mitogenome or COI_spike species, change column order
mock_data$mito_spike <- if_else(grepl("_COI$", mock_data$mitogenome_spike), "spike", "mitogenome")
mock_data <- mock_data %>% dplyr::select(sample, mito_spike, everything())
 
names(mock_data)

## save data files
# save(mock_inputspp, mock_design, mock_data, file = "~/Dropbox/Working_docs/Roslin_Greenland/2018/manuscript_methods/20190109/Rcode/Rdata/arcdyn_suppfiles_mocks_20190116_406barcodes.RData")

## load datafiles
# load("~/Dropbox/Working_docs/Roslin_Greenland/2018/manuscript_methods/20190109/Rcode/Rdata/arcdyn_suppfiles_mocks_20190116_406barcodes.RData")
 
```





